---
title: "p8105_hw1_sc4636"
output: github_document
---
# Homework 1

```{r setup, include = FALSE}
library(tidyverse)
```

## Problem 1

**Create a data frame**

```{r problem1 df}
problem1_df = tibble(
  vec_rnorm = rnorm(8),
  vec_logical = c(vec_rnorm > 0),
  vec_char = c("sky", "fire", "tomato", "grass", "ocean", "jeans", "cucumber", "strawberry"),
  vec_factor = factor(c("blue", "red", "red", "green", "blue", "blue", "green", "red"))
)
```

**Take the mean**

The variable `vec_rnorm` has a mean of ``r mean(pull(problem1_df, vec_rnorm))``, and the variable `vec_logical` has a mean of ``r mean(pull(problem1_df, vec_logical))``. 
Mean values can be taken directly from variables classified as _`r class(pull(problem1_df, vec_rnorm))`_ or _`r class(pull(problem1_df, vec_logical))`_.

Whereas, the mean for variable `vec_char` is ``r mean(pull(problem1_df, vec_char))`` and for `vec_factor` is ``r mean(pull(problem1_df, vec_factor))``. `mean()` cannot take means from variables classified as _`r class(pull(problem1_df, vec_char))`_ or _`r class(pull(problem1_df, vec_factor))`_.

**Convert types of variables**

The below chunk converts logical, character and factor variables in the above data frame to numeric data type.

```{r convert types, eval = FALSE}
as.numeric(pull(problem1_df, vec_logical))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))
```

When converting logical variables to numeric ones, `TRUE` logical variable takes numeric value of `1` while `FALSE` takes value of `0`. When taking the mean value of logical variables, R automatically converts logical variables to numeric ones. 

When converting character variables to numeric ones, in this case, a warning message `NAs introduced by coercion` popped. The character variables in this data frame are consisted of letters which would induce NA value when converted to numeric data type. Thus, when taking the mean value for character value in this data frame, the result shows `NA`.

When converting factor variables to numeric ones, R converts different factor levels in alphabetical order to match numeric variables if not specified. For example, factor level blue takes the value of 1, green takes 2, and red takes 3. However, the corresponding number has no meaning.

```{r manipulation of variables}
vec_logical_numeric = as.numeric(pull(problem1_df, vec_logical)) * pull(problem1_df, vec_rnorm)
vec_logical_factor = as.factor(pull(problem1_df, vec_logical)) * pull(problem1_df, vec_rnorm)
vec_logical_factor_numeric = as.numeric(vec_logical_factor) * pull(problem1_df, vec_rnorm)
```

## Problem 2

**Create a data frame**

```{r problem2 df}
problem2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_logical = c(x + y > 1),
  vec_numeric = as.numeric(vec_logical),
  vec_factor = as.factor(vec_logical)
)
```

* The above dataset is comprised of `r nrow(problem2_df)` rows and `r ncol(problem2_df)` columns.

* Variable `x` has a mean of `r mean(pull(problem2_df, x))`, a median of `r median(pull(problem2_df, x))`, and a standard deviation of `r sd(pull(problem2_df, x))`.

* The proportion of cases for which `x + y > 1` equals to 
`r mean(pull(problem2_df, vec_numeric))`.

**Making scatterplots**

```{r scatterplots of y vs x}
# Logical variable
ggplot(problem2_df, aes(x = x, y = y, color = vec_logical)) +
  geom_point() + labs(color = "Logical variable")
ggsave("plot1_logical_variable.png")

# Numeric variable
ggplot(problem2_df, aes(x = x, y = y, color = vec_numeric)) +
  geom_point() + labs(color = "Numeric variable")

# Factor variable
ggplot(problem2_df, aes(x = x, y = y, color = vec_factor)) +
  geom_point() + labs(color = "Factor variable")
```

These three plots show consistent distribution according to color. In the first and third plots, for dots in orange, `x + y ≤ 1`. For dots in blue, `x + y > 1`. In the second plot, `x + y > 1` for light blue dots while `x + y ≤ 1` for dark blue dots.